Template notebook for validating cell type assignments for an individual library in SCPCP000015

### 2025-01-29 analyses/cell-type-ewings/template_notebooks/celltype-exploration.Rmd
adjust plots based on review comments

all_info_df <- all_results_df |> 
  dplyr::left_join(gene_exp_df, by = "barcodes")
```

@@

# reformat auc data for density plots and UMAPs showing AUC values 
auc_df <- all_info_df |> 
  tidyr::pivot_longer(starts_with("auc_"), names_to = "geneset", values_to = "auc_value") |> 
  dplyr::mutate(
    geneset = stringr::str_remove(geneset, "auc_")
  ) |> 
  dplyr::select(barcodes, UMAP1, UMAP2, geneset, auc_value, cluster) |> 
  dplyr::left_join(auc_threshold_df, by = c("barcodes", "geneset")) |> 
  dplyr::mutate(
    auc_assignment =  auc_value > threshold
  )
```

@@

ggplot(auc_df, aes(x = auc_value, color = auc_assignment, fill = auc_assignment)) +
  geom_density(alpha = 0.5, bw = 0.01) +
  facet_wrap(vars(geneset)) +
  ggplot2::geom_vline(data = auc_df,
                      mapping = aes(xintercept = threshold),
                      lty = 2) +
  theme(
      aspect.ratio = 1,
      strip.background = element_rect(fill = "transparent", linewidth = 0.5),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5)
    )

```

@@

```{r}
ggplot(exp_df, aes(x = mean, color = tumor_cell_classification, fill = tumor_cell_classification)) +
  geom_density(alpha = 0.5, bw = 0.2) +
  facet_wrap(vars(geneset)) +
  theme(
    aspect.ratio = 1,
    strip.background = element_rect(fill = "transparent", linewidth = 0.5),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5)
  )
```

>>>>

all_info_df <- all_results_df |> 
  dplyr::left_join(gene_exp_df, by = "barcodes") |> 
  dplyr::arrange(cluster)
```

@@

# reformat auc data for density plots and UMAPs showing AUC values 
auc_df <- all_info_df |> 
  tidyr::pivot_longer(starts_with("auc_"), names_to = "geneset", values_to = "auc_value") |> 
  dplyr::mutate(
    geneset = stringr::str_remove(geneset, "auc_")
  ) |> 
  dplyr::select(barcodes, UMAP1, UMAP2, geneset, auc_value, cluster) |> 
  dplyr::left_join(auc_threshold_df, by = c("barcodes", "geneset")) |> 
  dplyr::mutate(
    in_geneset =  auc_value > threshold
  )
```

@@

ggplot(auc_df, aes(x = auc_value, color = in_geneset, fill = in_geneset)) +
  geom_density(alpha = 0.5, bw = 0.01) +
  facet_wrap(vars(geneset)) +
  ggplot2::geom_vline(data = auc_df,
                      mapping = aes(xintercept = threshold),
                      lty = 2) +
  theme(
      aspect.ratio = 1,
      strip.background = element_rect(fill = "transparent", linewidth = 0.5),
      panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5)
    ) 
```

@@

```{r, fig.height=8}
ggplot(exp_df, aes(x = mean, color = tumor_cell_classification, fill = tumor_cell_classification)) +
  geom_density(alpha = 0.5, bw = 0.2) +
  facet_wrap(vars(geneset)) +
  theme(
    aspect.ratio = 1,
    strip.background = element_rect(fill = "transparent", linewidth = 0.5),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5)
  )
```
